# Backpropagation & MLP

**Learning objectives:**

- Linear Models and ReLU
- Multi Layer Perceptron (MLP) from scratch

## Linear Models and ReLU {-}

```{r}
library(tidyverse)
theme_set(theme_bw())

x <- seq(-2, 2, by = 0.01) + 1
y <- x

gg_line <- ggplot() +
  geom_line(aes(x = x, y = y)) +
  ggtitle("Example linear model for y given x")
gg_line
```

Using only lines we want to draw a curve like the following:

```{r}
gg_line +
  geom_line(aes(x = x, y = (-(x-1)^2+2)), col = "blue")
```

- Can't add a bunch of lines because we will just get a line back

## Linear Models and ReLU {-}

What if we create piecewise lines like:

```{r}
gg_line +
  geom_line(aes(x = x, y = -0.5*ifelse(x>0, x, 0) - 1), col = "red")
```

Then add them together to get:

```{r}
re_lu1 <- x - 0.5*ifelse(x>0, x, 0) - 1

ggplot() +
  geom_line(aes(x = x, y = re_lu1)) +
  ggtitle("Example linear model for y given x")
```

Again:


```{r}
gg_line +
  geom_line(aes(x = x, y = -0.5*ifelse(x>0, x, 0) - 1), col = "red")+
  geom_line(aes(x = x, y = -0.5*ifelse(x-1>0, x-1, 0) - 1), col = "blue")
```

```{r}
re_lu2 <- x - 0.5*ifelse(x-1>0, x-1, 0) - 1

ggplot() +
  geom_line(aes(x = x, y = re_lu1), col = "red") +
  geom_line(aes(x = x, y = re_lu2), col = "blue") +
  ggtitle("Example linear model for y given x")
```

## Multi Layer Perceptron (MLP) from scratch {-}

Go to notebook

## Loss function from scratch - Mean Squared Error (MSE) {-}

Go to notebook; show screenshot

## Gradients and backpropagation diagram {-}

Reminders about SGD:

- move parameters (weights and biases) in (neg.) direction of gradients
- gradient matrix for every change in input what's the change in output (possibly more than 1)

## Matrix calculus resources {-}

[Matrix Calculus for Deep Learning](https://explained.ai/matrix-calculus/)

## Gradients and backpropagation code {-}

Go to notebook

## Chain rule visualized + how it applies {-}

[The Intuitive Notion of the Chain Rule](https://webspace.ship.edu/msrenault/geogebracalculus/derivative_intuitive_chain_rule.html)

Go through chain rule for $3x^2 + 9$?

## Using Python’s built in debugger {-}

- Add line to lin-grad
- press h for help
- explore shape of objects
- einsum demo to make change:

```python
w.g = (inp.unsqueeze(-1) * out.g.unsqueeze(1)).sum(0)
w.g = inp.T@out.g
```

## Refactoring the code {-}

- create a class for the functions
- __call__ : advantage to use it like a function, but instead is an instance of the class

## Minibatch Training {-}

- refresh on cross entropy with Excel file
- tricks with indexing vs one-hot-encoding
- refresh on log and exponent rules
- safe logsumexp by subtracting max

## Meeting Videos {-}

### Cohort 1 {-}

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>
<summary> Meeting chat log </summary>

```
LOG
```
</details>
